---
title: "nirmitin_OriginalHomeworkCode_04"
author: "Nirmiti Naik"
date: "10/20/2021"
output: html_document
---
<br>
The stringr package is installed so that the command str_split() will work. 
```{r}
library(stringr)
library(curl)
library(ggplot2)
library(manipulate)
library(gridExtra)
library(lmodel2)
```

## **Challenge 1**
*Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines:*
<br> 
My overall issue with this challenge is not really knowing how to fit all these subparts together. 
```{r}
Z.prop.test() <- function(p1, n1, p2 = NULL, n2 = NULL, p0, alternative = c("two.sided", "less", "greater"), conf.level = 0.95)
```

I'm struggling with what exactly the function should be executing in each of the if/else situations -> how do we code for p1 being tested rather than p2? 
**I am also a bit perplexed by that issue. However, I think that using the prop.test function might automatically work for both greater or less?**

```{r}
  #When conducting a two-sample test, it should be p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”, the same as in the use of x and y in the function t.test().
  if (alternative == "less") {
    if (p1 < p2) {
      mean <- mean(p1)
      sd <- sd(p1)
    }
  }

  if (alternative == "greater") {
    if (p1 < p2) {
      mean <- mean(p1)
      sd <- sd(p1)
    }
  }

#how to calculate variables if we're considering p1 and p2? <- p2 and n2 != null
mean <- 
ci <- 
p <- 
t<- 
```
**So, I think in order for a function to work you have to include all of the actions the function is doing within the {}. Otherwise, R doesn't know what your if else statements are being used for.**

I'm not sure if this makes sense since it's using the function itself - is this recursion? Would this even work? My intention was that by narrowing the scope of variables that can be used (i.e. only p1, n1, p0) the function would only take those inputs if p2 or n2 is null. 
```{r}
  #The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.
  if (is.null(p2) || is.null(n2)) {
    Z.prop.test(p1, n1, p0)
  } 

  if (!is.null(p2) || !is.null(n2))
    Z.prop.test(p1, n1, p0, p2, n2)
```

```{r}
  #The function should contain a check for the rules of thumb we have talked about (n∗p>5 and n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings. If this is violated, the function should still complete but it should also print an appropriate warning message.
  if (n∗p>5 & n∗(1−p)>5) {
    print("This isn't following a normal distribution")
  }
```
**Here there is an issue that the multiplication sign when copy and pasted from the R module is not recognized in R. I had to retype out this formula.**
```{r}
  #The function should return a list containing the members Z (the test statistic), P (the appropriate p value), and CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.
if #twosided
  Z
  P
  CI_p1

if #less/greater
  Z
  P
  CI_p1_p2
    
```
**The two if statements here don't work because they have a # infront of them which is normal reserved for comments. Also, they aren't following the typical syntax of an if statement. Again, I think your function will have issues running when split up into different R chunks like this.** 


## **Challenge 2**
<br>
*The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):* 
First, I will load the Kamilar and Cooper dataset.  
```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall21/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```
###longevity~brain size
Next, I will try to use the lm() function for longevity~brain size
```{r}
lm <- lm(MaxLongevity_m ~ Brain_Size_Species_Mean, data = d)
lm

names(lm)
lm$coefficients

head(lm$model)

g <- ggplot(data = d, aes(x = MaxLongevity_m, y = Brain_Size_Species_Mean))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g
```
**Looks good! I liked how you utilized lm$coefficients to look at just the coefficients of the model, I didn't know you could do that!**
###log(longevity)~log(brain size)

Now, I'll repeat the same code but for log(longevity)~log(brain size). It's interesting that this function removes rows containing "non-finite values and missing values). Visually, the confidence interval also looks wider for this graph. 
```{r}
loglm <- lm(log(MaxLongevity_m) ~ log(Brain_Size_Species_Mean), data = d)
loglm

names(loglm)
loglm$coefficients

head(loglm$model)

g <- ggplot(data = d, aes(x = log(MaxLongevity_m), y = log(Brain_Size_Species_Mean)))
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g
```
**It's interesting that you didn't have any issues logging the two variables, as the missing values gave me a hard time.**

We can make conclusions about the linear regression equations using summary()
```{r}
summary(lm)

ci <- confint(lm, level = 0.95)  # using the results of lm()
ci

summary(loglm)

ci <- confint(loglm, level = 0.95)  # using the results of lm()
ci
```
**Looks good! It might be useful to give the variables for the confidence intervals of the lm and loglm different names however, as this way they can be called individually**
Would we use code similar to this to plot the lines?
```{r}
v <- seq(from = 10, to = 30, by = 1)
m <- lm(data = d, height ~ age)
ci <- predict(m, newdata = data.frame(age = v), interval = "confidence", level = 0.95)
pi <- predict(m, newdata = data.frame(age = v), interval = "prediction", level = 0.95)
plot(data = d, height ~ age)
lines(x = v, y = ci[, 1], col = "black")
lines(x = v, y = ci[, 2], col = "blue")
lines(x = v, y = ci[, 3], col = "blue")
lines(x = v, y = pi[, 2], col = "red")
lines(x = v, y = pi[, 3], col = "red")
```
**I also had issues with this section. I think we are meant to use the predict function in order to plot these confidence intervals.**

**Overall, I think this was a great effort and was impressed by how you used novel solutions I would not have thought of. I think the main things to work on would be putting all of the if statements into the function in one chunk of code, and making sure the syntax of the if else statements are correct.**